Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-05-08 chapters/ch_experiment.tex
Added GG information

\section{Energy loss in material}
The average energy loss in a material can be described by the Bethe-Bloch equation,
\begin{equation}\label{eq:bb}
\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2).
\end{equation}
Where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle. There is a large variation in energy loss around this mean value, with a long high energy loss tail. A full description of the energy loss distribution can be described in CITE HERE.

\section{Time Projection Chambers}
%explain in breif the basics of a time projection chamber

@@

\section{S$\pi$RIT TPC Overview}
%Add Overview image with labels
The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (S$\pi$RI TPC) is a Multi-Wire Proportional Counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiments.  The TPC is enclosed in a thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach the arrays of scintillating bars detectors on the sides and downstream of the TPC. The S$\pi$RI TPC was developed to fit inside the Samurai dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC. More detail and specifications of the Samurai dipole magnet are given in \cite{samurai}.

A target ladder allowed for up to 5 fixed targets to be mounted. A ACME worm gear allowed for the x-axis motion for changing the targets during the experiment, without needing to open or move the TPC. The motion of this worm gear was translated through the target motion feed-through by several brass gears and non-magnetic gear boxes. The motion of the target ladder could be controlled by hand or by operation of the drill. The targets were mounted on stand-offs on the target ladder and also had z-axis motion. This allowed for the targets to be positioned as close as possible to the thin window of the field cage, maximizing the geometric acceptance.

The electronics were mounted to the aluminum top plate. Several aluminum ribs were mounted to the top of the plate to bring structural rigidity. A flatness within 150$\mu$m was achieved across the whole top plate as measured by a laser position system. The charge sensitive pads of the pad plane were etched into several circuit boards were recessed and glued to the bottom portion of the top plate. Vias through the pad plane circuit boards brought the signal traces from the pads to surface mount pads on the other side of the boards. Several holes cut through the top plate allowed for the interface cables of the electronics to be connected to these surface pads.

Just below the pad plane were a set of three wire planes; the gating grid, ground, and anode wire planes. The detailed function of these wires will be explained later but they served to separate the boundary between the drift volume and the avalanche volume.

The front and sides of the field cage were assembled from 8 independent rigid circuit boards. The downstream window served as the downstream wall of the field cage all though it was was a large polycarbonate frame in which a removable kapton window could be installed. This thin window minimized the scattering of exiting neutrons and charged particles for downstream detectors.

The voltage step down takes the high voltage of the  cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}rr@{}}\toprule 
\multicolumn{2}{c}{\spirit TPC Overview} \\
 \midrule
Pad plane area & 1.3 m x .9 m\\
Pad size       & 1.2 cm x .8 cm \\
Number of pads & 12096 (112 x 108) \\
Gas composition& 90\% Ar + 10\% CH${}_4$  \\
Multiplicity limit & 200  \\
dE/dx range        & Z=1-8, $\pi$, p,d,t,He,Li-O \\
Drift length       & 50 cm \\
\bottomrule
\end{tabular}
\caption{An overview of the properties of the \spirit TPC}
\label{tb:spiritoverview}
\end{table*}

\subsection{Voltage Step Down}
The gap between the field cage's cathode and the ground of the enclosure is quite small. To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor. There were 7 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt}.

ADD info how to calculate cathode voltage from HW supply.

The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires. It was designed to hang from the top plate and therefore needed to be of a lightweight construction. Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors. Therefore instead of a downstream wall, a large thin exit window was constructed. The cathode was constructed of an aluminum honeycomb laminate. Two sheets of ??? aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{fc_overview.pdf}
\label{fig:fc_overview}
\caption{FC overview}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{fc_overview2.pdf}
\label{fig:fc_overview2}
\caption{FC overview 2}
\end{figure}

The field cage was constructed from several panels of printed circuit boards (PCBs). The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires [CITE]. The halogen free material chosen was Cryogenic-G10. We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure. While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.

The front of the field cage was made of two PCBs and each side was constructed of three PCBs. The exit window was a 10$\mu$m thick Kapton window with evaporated aluminum strips on the inside and outside; mounted on a polycarbonate frame.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{FC_schematic.pdf}
\label{fig:FC_schematic}
\caption{Schematic of the electric connections relevant to the Field Cage system.}
\end{figure}
\subsection{Voltage Step Down}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{TPC_schematic.pdf}
\caption{Schematic of the TPC system}
\label{fig:TPC_schematic}
\end{figure}

Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls. The basic idea of the field cage is to set up a constant electric field in the field cage volume to allow for a constant drift, in the presence of gas, for electrons. To accomplish this the field cage is composed of a set of copper strips which are set at a defined voltage by a resistor chain for a fixed distance between strips; setting up a constant electric field.

The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail. The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor, $R$, to the next inside strip and the first outside strip. The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field. The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC. Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor with the option to place an additional resistor in parallel to tune the voltage of the top perimeter.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_open.pdf} 
        \caption{Wires open} \label{fig:wires_open}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_closed.pdf} 
        \caption{Wires closed.} \label{fig:wires_closed}
    \end{subfigure}
\label{fig:wires}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{gg_onoff.pdf}
\caption{On and off configurations of the gating grid.}
\label{fig:gg_onoff}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{gain.png}
\caption{Number of electrons produced in a single avalanche on an anode wire. Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$. The expected Polya distribution fit is also given in yellow.}
\label{fig:anodegain}
\end{figure}

@@

\subsection{Gas Properties}
The gas used inside of the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume --so called P10--  gas, operated just under atmospheric pressure (1 atm). The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere. The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants. The water never exceed  ??? ppm  and the oxygen level never exceeded ??? ppm. Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}. Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure. The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.

>>>>

\section{Operational Principles of Time Projection Chambers}

\begin{figure}
\includegraphics[width=\textwidth]{tpcPrinciple.pdf}
\label{fig:tpcPrinciple}
\caption{Operation principle of the TPC}
\end{figure}
%explain in breif the basics of a time projection chamber 
Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions. Here we will outline the physical principles involved in the TPC measurement. A detector gas is housed in a field cage, which also sets up a constant electric field. As charged particles pass through the gas, electron-ion pairs are separated from by passing tracks, the electrons are accelerated by the electric field and the ions move in the opposite direction. Since the mean free path of the electrons is very small they quickly collide with a gas molecule slowing down the electron which goes on to repeat the cycle. This microscopic behavior manifests into the electron appearing to drift at a constant velocity. These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field liberating more electron-ion pairs from the gas creating an avalanche. The avalanche electrons finally terminate either on the anode wire while the ions from the avalanche move slowly away from the anode wires creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.

Two of the 3 coordinates are determined from the charge distribution on the pad plane. The third dimension comes from projecting the electrons back in time utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics. The momentum information is extracted from measuring the radius of curvature of the track and the constant magnetic field value. The energy loss (<dE/dx>) is measured by a segmented charge sensitive pads on the pad-plane which are connected to the readout electronics. Particle identification is achieve through knowledge of the rigidity and the <dE/dx> of the tracks as will be discussed in latter sections. In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.

\section{S$\pi$RIT TPC Overview}
%Add Overview image with labels

The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiment.  The TPC is enclosure is an aluminum angle iron skelleton with thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC. The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC. More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.

A target mechanism allowed for up to 5 fixed targets to be mounted with the ability for changing the targets during the experiment without needing to open or move the TPC. The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate. The pad plane and wire plane structures are also mounted to the bottom of the top plate with the electronics are mounted on the top of the top plate. Several aluminum ribs were mounted to the top. Holes on the top plate allowed the electronics to be connected to the surface mount pads on the pad-plane.

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}rr@{}}\toprule 
\multicolumn{2}{c}{\spirit TPC Overview} \\
 \midrule
Pad plane area & 1.3 m x .9 m\\
Pad size       & 1.2 cm x .8 cm \\
Number of pads & 12096 (112 x 108) \\
Gas composition& 90\% Ar + 10\% CH${}_4$ (1 atm)  \\
Multiplicity limit & 200  \\
dE/dx range        & Z=1-3, $\pi$, p, d, t, He, Li \\
Drift length       & 50 cm \\
\bottomrule
\end{tabular}
\caption{An overview of the properties of the \spirit TPC}
\label{tb:spiritoverview}
\end{table*}

\subsection{Voltage Step Down}
The gap between the field cage's cathode and the ground of the enclosure is quite small. To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor. There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time. The first ring is the same voltage as the cathode and the last ring is connected to ground. All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.

The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires. It was designed to hang from the top plate and therefore needed to be of a lightweight construction. Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors. Therefore instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame with \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips. The cathode was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{fc_overview.png}
\label{fig:fc_overview}
\caption{FC overview}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{fc_overview2.png}
\label{fig:fc_overview2}
\caption{FC overview 2}
\end{figure}

The field cage was constructed from several panels of printed circuit boards (PCBs). The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires CITE HERE. The halogen free material chosen was Cryogenic-G10. We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure. While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only. The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.6]{FC_schematic.pdf}
\caption{Schematic of the electric connections relevant to the Field Cage system.}
\label{fig:FC_schematic}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{TPC_schematic.pdf}
\caption{Schematic of the TPC system}
\label{fig:TPC_schematic}
\end{figure}

The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated from the schematic of the TPC system in Fig.~\ref{fig:TPC_schematic}, as

\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( \frac{1}{245 + R_{eff}} + \frac{1}{700} \right)^{-1} } },
\end{equation}

where $R_{eff} = \left( 1/20 + 1/R_{TP2} \right)^{-1}$, and $V_{HV}$ is the high voltage supply; all resistor values are given in \si{\mega\ohm}.

Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls. The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail. The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the next inside strip and the first outside strip. The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field. The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC. Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{gain.png}
\caption{Number of electrons produced in a single avalanche on an anode wire. Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$. The expected Polya distribution fit is also given in yellow.}
\label{fig:anodegain}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_open.pdf} 
        \caption{Wires open} \label{fig:wires_open}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.42\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_closed.pdf} 
        \caption{Wires closed.} \label{fig:wires_closed}
    \end{subfigure}
\label{fig:wires}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{gg_onoff.pdf}
\caption{On and off configurations of the gating grid.}
\label{fig:gg_onoff}
\end{figure}

\section{Energy loss in material}
The average energy loss in a material can be described by the Bethe-Bloch equation,
\begin{equation}\label{eq:bb}
\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2).
\end{equation}
Where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle. There is a large variation in energy loss around this mean value, with a long high energy loss tail.

The statistical variation of energy loss in a material was described by Landau \cite{landau} and later better described by Shulek \cite{shulek} and Bichsel \cite{bichsel1}. In both approximations it is described by a most probable energy loss value with a long, high-energy loss tail. Because of this long tail, for a finite set of energy loss measurements along a given track, the fluctuation of the mean value energy loss is a very unreliable observable. The most probable energy loss is the most desirable observable either through fitting of the observed distribution or through the truncated mean method. The truncated mean is the average mean value calculated after throwing away the top fraction of the highest energy loss entries, where the fraction used in this thesis was 30\%. This approximates the most probable value without performing a fit to a known distribution. There are

%Suppose $E$ represents the set of $N$ energy loss points measured, sorted from lowest to highest energy loss. The $j^{th}$ index marks the position where  $i > j$ entries represent the highest value of energy loss measurements, where their fraction of the total set is expressed as, $f_c = \sum_{i>j} E_i/ \sum_i E_i$. The truncated mean is expressed as the mean value of the remaining energy loss measurements, throwing away the top $f_c$ fraction, expressed as,

%\begin{equation}
%<dE/dx>_t = \frac{\sum_{i < j} E_i}{N}
%\label{eq:truncatedM}
%\end{equation}

 A full description of the energy loss distribution can be described in CITE HERE.

\subsection{Gas Properties}
The gas contained by the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure (1 atm). The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere. The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants. The water never exceed  ??? ppm  and the oxygen level never exceeded ??? ppm. Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}. Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure. The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.

@@

@@
